---
title: "StressAware Heart Rate Data Analysis Report"
authors: "Paula Banach (440186), Igor Kołodziej (440239)"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[L]{StressAware Analysis}
- \fancyhead[R]{\thepage}
- \fancyfoot[C]{}
params: # All parameters must be declared here (leave this as 'params')
  hr_series_obj_param: NULL
  classified_stress_data_param: NULL
  plot_types_param: NULL
  project_root_param: "" # Declare the new parameter
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4)

# Load required libraries for the R Markdown document
library(methods)
library(ggplot2)
library(dplyr)
library(patchwork)
library(lubridate)
library(hrbrthemes)
library(readr)
library(zoo)
library(reticulate)
library(knitr)
library(tinytex) # For PDF output if using TinyTeX

# Access the parameters passed from the Shiny app using the NEW variable name
# The 'params' in the YAML header (above) is still 'params', but the object
# actually passed into the Rmd's environment via the `rmarkdown::render(params=...)`
# argument now has a different name, which will become a variable in this chunk.
report_params <- params # 'params' here will actually contain the list passed as 'report_params'

# Access all parameters passed from the Shiny app using the new variable name
project_root <- report_params$project_root_param
hr_series_obj <- report_params$hr_series_obj_param
classified_stress_data <- report_params$classified_stress_data_param
plot_types <- report_params$plot_types_param


# --- Debug messages (can remove after verification) ---
message("DEBUG (Rmd setup): Current working directory for Rmd: ", getwd())
message("DEBUG (Rmd setup): Project root passed from server: ", project_root)

# Construct the absolute path to your 'R/' directory
# Assuming 'R/' is directly under the project root
r_script_path <- file.path(project_root, "R")
message("DEBUG (Rmd setup): R scripts directory: ", r_script_path)


# Ensure the Python environment is configured if not done in server.R's global environment
# It's highly recommended to do `reticulate::use_python()` or `reticulate::use_condaenv()`
# in your `server.R` or `global.R` file so it's set up before the Rmd renders.
# However, for robustness, you can add a fallback here if absolutely necessary,
# but note that `rmarkdown::render` runs in a clean environment by default.
# The `envir = new.env(parent = globalenv())` in server.R helps, but re-sourcing
# Python configuration might be needed if functions relying on reticulate are called
# directly within the Rmd and not just passed objects.
tryCatch({
  if (reticulate::py_available()) {
    message("Reticulate Python is available.")
  } else {
    warning("Reticulate Python is not available. Please check configuration in server.R.")
  }
  reticulate::py_discover_config()
}, error = function(e) {
  warning("Reticulate Python environment configuration failed in Rmd: ", e$message,
          " Ensure Python with 'joblib' and 'numpy' is installed and accessible.")
})


# Load the Python stress model globally if your Rmd needs to call model functions
# Adjust path if your model is elsewhere. If model/ is parallel to app/
# and project_root is the parent of app/, then model_file_path would be project_root/model/stress_rf.pkl
model_file_path <- file.path(project_root, "stress_rf.pkl") # Assuming 'stress_rf.pkl' is at the project root

if (file.exists(model_file_path)) {
  # It's crucial that `load_stress_model` and `classify_stress_windows` functions
  # (from model.R) are available in the Rmd's rendering environment.
  # If server.R sources model.R and passes the global environment, they should be.
  # If you are knitting this Rmd file directly for testing, you might need to source:
  # source(file.path(r_script_path, "model.R"))
  message("Python stress model expected at: ", model_file_path)

  tryCatch({
    if (exists("load_stress_model", mode = "function")) {
       load_stress_model(model_path = model_file_path)
       message("Python stress model successfully loaded (or re-loaded) for Rmd use.")
    } else {
       warning("`load_stress_model` function not found in Rmd environment. Cannot explicitly load model.")
    }
  }, error = function(e) {
    warning("Failed to load Python stress model for Rmd: ", e$message)
  })
} else {
  warning(paste0("Stress classification model '", model_file_path, "' not found for Rmd. ",
                 "Stress classification features will not be available in report. Please ensure the model file is at your project root."))
}


# Ensure the hr_series_obj is created or handled if not passed from Shiny
# For stand-alone knitting (e.g., direct "Knit" button), you might need to load sample data here
# This block is for Rmd knitting when not run from Shiny, to provide demo data.
if (is.null(hr_series_obj) || is.null(classified_stress_data) || nrow(classified_stress_data) == 0) {
  message("HRSeries object or classified stress data not provided as parameter. Loading sample data for Rmd knitting demo.")
  hr_data_dir <- file.path(project_root, "hr_data")
  sample_file_path <- file.path(hr_data_dir, "3.txt")

  tryCatch({
    source(file.path(r_script_path, "classes.R"))
    source(file.path(r_script_path, "clean_hr_initial.R"))
    source(file.path(r_script_path, "detect_artifacts_malik.R"))
    source(file.path(r_script_path, "interpolate_hr.R"))
    source(file.path(r_script_path, "preprocess_hr.R"))
    source(file.path(r_script_path, "extract_features.R"))
    source(file.path(r_script_path, "model.R"))
    message("Successfully sourced R scripts for Rmd demo data generation.")
  }, error = function(e) {
    warning("Failed to source R scripts for Rmd demo data generation: ", e$message)
    hr_series_obj <- NULL
    classified_stress_data <- NULL
  })

  if (file.exists(sample_file_path) && exists("preprocess_hr") && exists("extract_hrv_features") && exists("create_hrseries")) {
      tryCatch({
        raw_data_demo <- readr::read_csv(sample_file_path, col_types = readr::cols(.default = "c"), show_col_types = FALSE)
        preproc_results_demo <- preprocess_hr(
          data = raw_data_demo,
          time_col = "Time",
          rr_col = "rr_ms",
          resample_rate = 4, 
          malik_threshold = 0.2
        )
        hrv_features_demo <- extract_hrv_features(
          preprocessed_results = preproc_results_demo,
          fs = 4,
          window_seconds = 120,
          step_seconds = 60
        )
        hr_series_obj <- create_hrseries(
          preprocessed_data = preproc_results_demo,
          features_data = hrv_features_demo,
          file_source = basename(sample_file_path),
          subject_id = "DemoSubject",
          preprocessing_params = list(resample_rate = 4, malik_threshold = 0.2),
          feature_extraction_params = list(fs = 4, window_seconds = 120, step_seconds = 60)
        )

        if (!is.null(hrv_features_demo) && nrow(hrv_features_demo) > 0 && exists("classify_stress_windows")) {
          classified_stress_data <- classify_stress_windows(hrv_features_df = hrv_features_demo, model_path = model_file_path)
          message("Successfully classified stress for Rmd demo data.")
        } else {
          classified_stress_data <- NULL
          warning("No HRV features extracted or classify_stress_windows not found for Rmd demo, cannot classify stress.")
        }

      }, error = function(e) {
        warning("Error loading or processing sample data for Rmd demo: ", e$message)
        hr_series_obj <- NULL
        classified_stress_data <- NULL
      })
  } else {
    warning("Could not load sample data or find necessary processing functions for Rmd demo.")
  }
}

if (is.null(plot_types) || length(plot_types) == 0) {
  plot_types <- c("cleaned_rr", "interpolated_hr", "hrv_features", "stress_classification")
  message("DEBUG (Rmd setup): plot_types was NULL or empty, setting to default for Rmd demo.")
} else {
  message("DEBUG (Rmd setup): plot_types received: ", paste(plot_types, collapse = ", "))
}

data_available <- !is.null(hr_series_obj) && !is.null(hr_series_obj@cleaned_rr) && nrow(hr_series_obj@cleaned_rr) > 0
```
# Introduction 

This report presents an analysis of heart rate (HR) and heart rate variability (HRV) data, including preprocessing steps, feature extraction, and stress classification using a pre-trained machine learning model.

# Data Summary and Preprocessing

This section provides an overview of the raw data and the effects of the preprocessing pipeline.

```{r data-summary, echo=FALSE}
if (data_available) {
  cat("### Basic Data Characteristics\n")
  cat(paste0("- **File Source:** ", hr_series_obj@metadata$file_source, "\n"))
  cat(paste0("- **Subject ID:** ", hr_series_obj@metadata$subject_id, "\n"))
  cat(paste0("- **Processing Timestamp:** ", hr_series_obj@metadata$processing_timestamp, "\n"))
  cat(paste0("- **Original Data Points:** ", nrow(hr_series_obj@metadata$original_data), "\n"))
  cat(paste0("- **Cleaned RR Intervals (Initial QC Passed):** ", sum(hr_series_obj@cleaned_rr$initial_qc_passed), " / ", nrow(hr_series_obj@cleaned_rr), "\n"))
  cat(paste0("- **Artifacts Detected (Malik Criterion):** ", sum(hr_series_obj@cleaned_rr$is_artifact), " / ", nrow(hr_series_obj@cleaned_rr), "\n"))
  cat(paste0("- **Interpolated HR Points:** ", nrow(hr_series_obj@interpolated_hr), "\n"))
  cat(paste0("- **Resampling Rate:** ", hr_series_obj@metadata$preprocessing_params$resample_rate, " Hz\n"))
} else {
  cat("No data available to generate summary and plots. Please ensure data is correctly loaded.\n")
}
```

## Cleaned RR Intervals

This plot shows the raw and cleaned RR intervals, highlighting points identified as artifacts and those that failed initial quality control.

```{r cleaned-intervals, echo=FALSE}
if (data_available && "cleaned_rr" %in% plot_types) {
  # Plot using the method defined in classes.R (or plot_preprocessing_results if available)
  # Assuming plot method of HRSeries object is available and handles "cleaned_rr" type
  plot(hr_series_obj, type = "cleaned_rr",
       title = "Raw and Cleaned RR Intervals with Artifacts")
}
```

## Interpolated Heart Rate

This plot displays the instantaneous heart rate after interpolation over artifacts and resampling to a uniform time grid.

```{r interpolated-HR, echo=FALSE}
if (data_available && "interpolated_hr" %in% plot_types) {
  plot(hr_series_obj, type = "interpolated_hr",
       title = "Interpolated and Resampled Heart Rate")
}
```

# Heart Rate Variability (HRV) Features

This section presents the extracted Time-Domain and Frequency-Domain HRV features over defined windows.

```{r HRV-features, echo=FALSE}
if (data_available && !is.null(hr_series_obj@hrv_features) && nrow(hr_series_obj@hrv_features) > 0) {
  cat("### Extracted HRV Features (First 5 Windows)\n")
  print(head(hr_series_obj@hrv_features))
  cat("\n### HRV Feature Summary Statistics\n")
  print(summary(dplyr::select(hr_series_obj@hrv_features, sdnn, rmssd, lf_power, hf_power, lf_hf_ratio)))
} else {
  cat("No HRV features available.\n")
}
```

## HRV Features Over Time

Plots showing the change in key HRV features (e.g., RMSSD, LF/HF Ratio) over the recording period.

```{r HRV-overtime, echo=FALSE}
if (data_available && "hrv_features" %in% plot_types && !is.null(hr_series_obj@hrv_features) && nrow(hr_series_obj@hrv_features) > 0) {
  plot(hr_series_obj, type = "hrv_features",
       title = "HRV Features Over Time")
}
```

# Stress Classification

This section provides the results of stress classification based on HRV features, using a pre-trained machine learning model.

```{r stress-classif, echo=FALSE}
if (data_available && !is.null(classified_stress_data) && nrow(classified_stress_data) > 0) {
  cat("### Stress Classification Results (First 5 Windows)\n")
  print(head(classified_stress_data))
  cat("\n### Stress Prediction Summary\n")
  print(table(classified_stress_data$stress_prediction, useNA = "ifany"))
} else {
  cat("No stress classification results available.\n")
}
```

## Stress Classification Over Time

This plot visualizes the stress predictions over time, often alongside the associated probability.

```{r stress-overtime, echo=FALSE}
if (data_available && "stress_classification" %in% plot_types && !is.null(classified_stress_data) && nrow(classified_stress_data) > 0) {
  
  plot_data_rmd <- classified_stress_data %>%
    mutate(
      stress_level_label = case_when(
        stress_prediction == 1 ~ "High Stress",
        stress_prediction == 0 ~ "Low Stress",
        TRUE ~ "Unknown" # Should not happen if predictions are only 0 or 1
      ),
      stress_level_label = factor(stress_level_label, levels = c("Low Stress", "High Stress", "Unknown"))
    )

  p_stress <- ggplot(plot_data_rmd, aes(x = window_end_time, y = stress_prediction, color = stress_level_label)) +
    geom_point(size = 2.5, alpha = 0.8) + 
    scale_color_manual(values = c("Low Stress" = "forestgreen", "High Stress" = "firebrick", "Unknown" = "grey50"), 
                       name = "Stress Level") +
    labs(title = "Stress Prediction Over Time",
         x = "Time",
         y = "Stress Prediction") +
    scale_y_continuous(breaks = c(0, 1), labels = c("Low (0)", "High (1)"), limits = c(-0.1, 1.1)) +
    theme_minimal(base_size = 10) + 
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = 45, hjust = 1))

  print(p_stress)
} else {
   ggplot() + theme_void() + labs(title = "No stress classification data to plot.") +
   annotate("text", x = 0.5, y = 0.5, label = "No stress classification data to display.", size = 5)
}
```
